Unicode Implementation Issues \thanks { This is a working document for the Rexx Language Association (RexxLA). There will be one or more separate requirements documents. It is written in \LaTeX\ \fmtversion\ and was rendered \today\ using \enginedetails. }

### 2024-04-12 ARB/standards/work-in-progress/document/tex/standard/built-in_functions.tex
updates

/* Check arguments. Some further checks will be made in particular
built-ins.\emph{/ /} The argument to CheckArgs is a checklist for the
allowable arguments. */

/* NUM, WHOLENUM and WHOLE have a side-effect, `normalizing' the number.
*/

/* Calls to raise syntax conditions will not return. */

CheckArgs: CheckList = arg(1) /* This refers to the argument of
CheckArgs. */

/* Move the checklist information from a string to individual variables
\emph{/ ArgType. = '\,'\,' ArgPos = 0 /} To count arguments \emph{/
MinArgs = 0 do j = 1 to length (CheckList) ArgPos = ArgPos+1 /} Count
the required arguments. \emph{/ if substr(CheckList,j,1) == `r' then
MinArgs = MinArgs + 1 /} Collect type information. \emph{/ do while j
\textless{} length(CheckList) j=eajae1 t = substr(CheckList,j,1) if t=='
' then leave ArgType.ArgPos = ArgType.ArgPos \textbar\textbar{} t end /}
A single space delimits parts. */ end j MaxArgs = ArgPos

/* Check the number of arguments to the built-in, in this instance. */
NumArgs = \#Bif Arg.0

if NumArgs \textless{} MinArgs then call Raise 40.3, MinArgs

if NumArgs \textgreater{} MaxArgs then call Raise 40.4, MaxArgs

/* Check the type(s) of the arguments to the built-in. */ do ArgPos = 1
to NumArgs if \#Bif ArgExists.ArgPos then call CheckType else if ArgPos
\textless= MinArgs then call Raise 40.5, ArgPos end ArgPos

/* No errors found by CheckArgs. */

return CheckType: Value = \#Bif Arg.ArgPos Type = ArgType.ArgPos select
when Type == `ANY' then nop /* Any string \emph{/ when Type == `NUM'
then do /} Any number */

/* This check is made with the caller's digits setting. \emph{/ if
\Cdatatype(Value, `N') then if \#DatatypeResult==`E' then call Raise
40.9, ArgPos, Value else call Raise 40.11, ArgPos, Value \#Bif
Arg.ArgPos=\#DatatypeResult /} Update argument copy. */ end

when Type == `WHOLE' then do /* Whole number \emph{/ /} This check is
made with digits setting for the built-in. */ if \Edatatype(Value,`W')
then call Raise 40.12, ArgPos, Value \#Bif\_ Arg.ArgPos=\#DatatypeResult
end

when Type == `WHOLE\textgreater=0' then do /* Non-negative whole number
*/ if \Edatatype(Value,`W') then call Raise 40.12, ArgPos, Value if
\#DatatypeResult \textless{} 0 then call Raise 40.13, ArgPos, Value
\#Bif\_ Arg.ArgPos=\#DatatypeResult end

when Type == `WHOLE\textgreater0O' then do /* Positive whole number */
if \Edatatype(Value,`W') then call Raise 40.12, ArgPos, Value if
\#DatatypeResult \textless= 0 then call Raise 40.14, ArgPos, Value
\#Bif\_ Arg.ArgPos=\#DatatypeResult end

when Type == `WHOLENUM' then do /* D2X type whole number \emph{/ /} This
check is made with digits setting of the caller. */ if
\Cdatatype(Value,`W') then call Raise 40.12, ArgPos, Value \#Bif\_
Arg.ArgPos=\#DatatypeResult end

when Type == `WHOLENUM\textgreater=0' then do /* D2X Non-negative whole
number */ if \Cdatatype(Value,`W') then call Raise 40.12, ArgPos, Value
if \#DatatypeResult \textless{} 0 then call Raise 40.13, ArgPos, Value
\#Bif\_ Arg.ArgPos=\#DatatypeResult end

when Type == `0 90' then do /* Errortext */

if \Edatatype(Value,`N') then call Raise 40.11, ArgPos, Value

Value=\#DatatypeResult

\#Bif\_ Arg.ArgPos=Value

Major=Value \% 1

Minor=Value - Major

if Major \textless{} 0 \textbar{} Major \textgreater{} 90 \textbar{}
Minor \textgreater{} .9 \textbar{} pos(`E',Value)\textgreater0 then call
Raise 40.17, Value /* ArgPos will be 1 */

end

when Type == `PAD' then do /* Single character, usually a pad. */

if length(Value) = 1 then call Raise 40.23, ArgPos, Value end

when Type == `HEX' then /* Hexadecimal string \emph{/ if
\datatype (Value, `X') then call Raise 40.25, Value /} ArgPos will be 1
\emph{/ when Type == `BIN' then /} Binary string \emph{/ if
\datatype(Value,`B') then call Raise 40.24, Value /} ArgPos will be 1
\emph{/ when Type == `SYM' then /} Symbol */

if \datatype (Value, `S') then call Raise 40.26, Value /* ArgPos will be
1 */

when Type == `STREAM' then do call Config Stream Qualify Value if left
(\#Response, 1) == `B' then call Raise 40.27, Value /* ArgPos will be 1
*/ end

when Type = `ACEFILNOR' then do /* Trace \emph{/ Val = Value /} Allow
`?' alone */

if val == `?' then do /* Allow leading `?' */ if left(Val,1) == `?' then
Val = substr(Val,2)

if pos(translate(left(Val, 1)), `ACEFILNOR') = 0 then call Raise 40.28,
ArgPos, Type, Val

end end otherwise do /* Options \emph{/ /} The checklist item is a list
of allowed characters */ if Value == '\,' then

call Raise 40.21, ArgPos \#Bif Arg.ArgPos = translate(left(Value, 1)) if
pos(\#Bif Arg.ArgPos, Type) = 0 then call Raise 40.28, ArgPos, Type,
Value end

end /* Select */ return

Cdatatype: /* This check is made with the digits setting of the caller.
\emph{/ /} \#DatatypeResult will be set by use of datatype() */

numeric digits \#Digits.\#Level

numeric form value \#Form.\#Level

return datatype(arg(1), arg(2))

Edatatype: /* This check is made with digits setting for the particular
built-in. \emph{/ /} \#DatatypeResult will be set by use of datatype()
*/

numeric digits \#Bif Digits.\#Bif

numeric form scientific

return datatype(arg(1),arg(2))

10.2.2 Date calculations Time2Date: if arg(1) \textless{} 0 then call
Raise 40.18 if arg(1) \textgreater= 315537897600000000 then call Raise
40.18 return Time2Date2 (arg(1))

Time: procedure /* This routine is essentially the code from the
standard, put in stand-alone form. The only `tricky bit' is that there
is no Rexx way for it to fail with the same error codes as a ``real''
implementation would. It can however give a SYNTAX error, albeit not the
desirable one. This causing of an error is done by returning with no
value. Since the routine will have been called as a function, this
produces an error. */

/* Backslash is avoided as some systems don't handle that negation sign.
\emph{/ if argQ)\textgreater3 then returm numeric digits 18 if
arg(1,`E') then if pos(translate(left(arg(1),1)),``CEHLMNRS'')=0 then
return /} (The standard would also allow `O' but what this code is
running on would not.) \emph{/ if arg(3,`E') then if
pos(translate(left(arg(3),1)),``CHLMNS'')=0 then return /} Tf the third
argument is given then the second is mandatory. \emph{/ if arg(3,`E') \&
arg(2,`E')=0 then returm /} Default the first argument. \emph{/ if
arg(1,`E') then Option = translate(left(arg(1),1)) else Option = `N'' /}
Tf there is no second argument, the current time is returned. \emph{/ if
arg(2,`E') = 0 then if arg(1,`E') then return ``TIME'(arg(1)) else
return `TIME'() /} One cannot convert to elapsed times. */ if
pos(Option, `ERO') \textgreater{} 0 then returm InValue = arg(2) if
arg(3,`E') then InOption = arg(3) else InOption = `N' HH =0 MM = 0 SS =0
HourAdjust = 0 select

when InOption == `C' then do parse var InValue HH `:'. +1 MM +2 XX if HH
= 12 then HH =0 if XX == `pm' then HourAdjust = 12

end when InOption == `H' then HH = InValue

when InOption == `L'' \textbar{} InOption == `N' then parse var InValue
HH `:' MM `:' SS when InOption == `M' then MM = InValue otherwise SS =
InValue end if datatype(HH,`W')=0 \textbar{} datatype(MM,`W')=0
\textbar{} datatype(SS,`N')=0 then returm HH = HH + HourAdjust /*
Convert to microseconds \emph{/ Micro = trunc((((HH } 60) + MM) * 60 +
SS) * 1000000) /* There is no special message for time-out-of-range; the
bad-format message is used. \emph{/ if Micro\textless0O \textbar{} Micro
\textgreater{} 24}3600* 1000000 then returm /* Reconvert to further
check the original. */ if TimeFormat(Micro,InOption) == InValue then
return TimeFormat(Micro, Option) return

TimeFormat: procedure /* Convert from microseconds to given format.
\emph{/ /} The day will be irrelevant; actually it will be the first day
possible. */ x = Time2Date2(arg(1)) parse value x with Year Month Day
Hour Minute Second Microsecond Base Days select when arg(2) == `C' then
select when Hour\textgreater12 then return
Hour-12':`right(Minute,2,'0'')'pm' when Hour=12 then return
`12:'right(Minute,2,'0')`pm' when Hour\textgreater0 then return
Hour':`right(Minute,2,'0')`am' when Hour=0 then return
`12:'right(Minute,2,'0')`am'

when arg(2) == `H' then return Hour when arg(2) == `L' then return
right(Hour,?2,`0')`:'right(Minute,2,'0')`:'right(Second,2,'0'),
\textbar\textbar{} `.'right(Microsecond,6,'0') when arg(2) == `M' then
return 60\emph{Hour+Minute when arg(2) == `N' then return
right(Hour,?2,`0')`:'right(Minute,2,'0')`:'right(Second,2,'0') otherwise
/} arg(2) ==`S' \emph{/ return 3600}Hour+60* Minute+Second end

Time2Date2: Procedure /* Convert a timestamp to a date. Argument is a
timestamp (the number of microseconds relative to 0001 01 01
00:00:00.000000) Returns a date in the form: year month day hour minute
second microsecond base days */

/* Argument is relative to the virtual date 0001 01 01 00:00:00.000000
*/ Time = arg(1)

Second = Time \% 1000000 \} Microsecond = Time // 1000000 Minute =
Second \% 60 ; Second = Second // 60 Hour = Minute \% 60 ; Minute =
Minute // 60 Day = Hour \% 24 ;\} Hour = Hour // 24

/* At this point, the days are the days since the 0001 base date. */
BaseDays = Day Day = Day + 1

/* Compute either the fitting year, or some year not too far earlier.
Compute the number of days left on the first of January of this year. */

Year = Day \% 366 Day = Day - (Year*365 + Year\%4 - Year\%100 +
Year\%400) Year = Year +1

/* Now if the number of days left is larger than the number of days in
the year we computed, increment the year, and decrement the number of
days accordingly. */ do while Day \textgreater{} (365 + Leap (Year) )

Day = Day - (365 + Leap(Year) )

Year = Year + 1 end

/* At this point, the days left pertain to this year. */ YearDays = Day

/* Now step through the months, increment the number of the month, and
decrement the number of days accordingly (taking into consideration that
in a leap year February has 29 days), until further reducing the number
of days and incrementing the month would lead to a negative number of
days */ Days = `31 28 31 30 31 30 31 31 30 31 30 31' do Month = 1 to
words (Days)

ThisMonth = Word(Days, Month) + (Month = 2) * Leap (Year)

if Day \textless= ThisMonth then leave

Day = Day - ThisMonth end

return Year Month Day Hour Minute Second Microsecond BaseDays YearDays

Leap: procedure /* Return 1 if the year given as argument is a leap
year, or 0 otherwise. */ return (arg(1)//4 = 0) \& ((arg(1)//100
\textless\textgreater{} 0) \textbar{} (arg(1)//400 = 0))

@@

ReRadix: /* Converts Arg(1) from radix Arg(2) to radix Arg(3) */

procedure

Subj ect=arg(1)

FromRadix=arg (2)

ToRadix=arg (3)

/* Radix range is 2-16. Conversion is via decimal */

Integer=0

do j=1 to length (Subject) /* Individual digits have already been
checked for range. \emph{/ Integer=Integer}FromRadix+pos (substr
(Subject,j,1),`0123456789ABCDEF')-1 end

rete

do while Integer\textgreater0

r= substr(`0123456789ABCDEF',1 + Integer // ToRadix, 1)
\textbar\textbar{} r Integer = Integer \% ToRadix end

/* When between 2 and 16, there is no zero suppression. \emph{/ if
FromRadix = 2 \& ToRadix = 16 then r=eright(r, (length(Subject)+3) \% 4,
`0') else if FromRadix = 16 \& ToRadix = 2 then reright(r,
length(Subject) } 4, `0') return r

@@

Raise:

/* These 40.nn messages always include the built-in name as an
insert.\emph{/ call \#Raise `SYNTAX', arg(1), \#Bif, arg(2), arg(3),
arg(4) /} \#Raise does not return. */

@@

ABBREV returns `1' if the second argument is equal to the leading
characters of the first and the length of the second argument is not
less than the third argument.

call CheckArgs `rANY rANY oWHOLE\textgreater=0'

Subject \#Bif Arg.1

Subj \#Bif Arg.2

if \#Bif\_ArgExists.3 then Length = \#Bif Arg.3 else Length = length
(Subj)

Condl = length(Subject) \textgreater= length(Subj) Cond2 = length(Subj)
\textgreater= Length Cond3 = substr(Subject, 1, length(Subj)) == Subj

return Condl \& Cond2 \& Cond3

@@

call CheckArgs `rANY rWHOLE\textgreater=0 oPAD'

String = \#Bif Arg.1

Length = \#Bif Arg.2

if \#Bif\_ArgExists.3 then Pad = \#Bif Arg.3 else Pad = ' '!

Trim = length(String) - Length

if Trim \textgreater{} 0 then return substr(String, Trim \% 2 + 1,
Length)

return overlay(String, copies(Pad, Length), -Trim \% 2 + 1)

@@

call CheckArgs ```rANY rANY rANY'

Output = '\,' Position = 1 do forever FoundPos = pos(\#Bif Arg.1, \#Bif
Arg.2, Position) if FoundPos = 0 then leave Output = Output
\textbar\textbar{} substr(\#Bif\_ Arg.2, Position, FoundPos - Position),
\textbar\textbar{} \#Bif Arg.3 Position = FoundPos + length(\#Bif Arg.1)
end return Output \textbar\textbar{} substr(\#Bif Arg.2, Position)

@@

call CheckArgs `rANY rANY oPAD'

Strl = \#Bif\_Arg.1

Str2 = \#Bif\_Arg.2

if \#Bif\_ArgExists.3 then Pad else Pad

\#Bif Arg.3

/* Compare the strings from left to right one character at a time */ if
length(Str1) \textgreater{} length(Str2) then do

Length = length(Str1)

Str2=left (Str2, Length, Pad)

end else do

Length = length(Str2)

Strl=left (Str1, Length, Pad)

end

do i= 1 to Length if substr(Stri, i, 1) == substr(Str2, i, 1) then
return i end

return 0

@@

call CheckArgs ```rANY rWHOLE\textgreater=0' Output = '\,' do \#Bif
Arg.2 Output = Output \textbar\textbar{} \#Bif\_Arg.1 end

return Output

@@

call CheckArgs ```rANY rANY'

Output = 0

Position = pos (\#Bif Arg.1,\#Bif Arg.2)

do while Position \textgreater{} 0 Output = Output + 1 Position =
pos(\#Bif Arg.1, \#Bif Arg.2, Position + length(\#Bif Arg.1)) end

return Output

@@

call CheckArgs `rANY oABLMNSUWX'

/* As well as returning the type, the value for a `NUM' is set in
\#DatatypeResult. This is a convenience when DATATYPE is used

by CHECKARGS. */

String = \#Bif Arg.1

/* If no second argument, DATATYPE checks whether the first is a number.
*/ if \#Bif\_ArgExists.2 then return DtypeOne()

Type = \#Bif\_ Arg.2 /* Null strings are a special case. */

if String == '\,' then do if Type == ``X'' then return 1 if Type ==
``B'' then return 1 return 0 end

/* Several of the options are shorthands for VERIFY \emph{/
azl=``abcdefghijklmnopqrstuvwxyz'' AZU= ``ABCDEFGHIJKLMNOPQRSTUVWXYZ''
DO09=``0123456789'' if Type == ``A'' then return
verify(String,az1\textbar\textbar AzU\textbar\textbar D09)=0 if Type ==
``B'' then do /} Check blanks in allowed places. \emph{/ if pos (left
(String,1),\#Al1Blanks)\textgreater0 then return 0 if pos (right
(String,1),\#Al1Blanks)\textgreater0 then return 0 BinaryDigits=0 do j =
length(String) by -1 tol ce = substr(String,j,1) if
pos(c,\#Al11Blanks)\textgreater0 then do /} Blanks need four
BinaryDigits to the right of them. */ if BinaryDigits//4 = 0 then return
0 end else do if verify(c,``01'') = 0 then return 0 BinaryDigits =
BinaryDigits + 1

end end j

return 1

end /* B */ if Type == ``L'' then return (verify (String,azl1)=0) if
Type == ``M'' then return (verify (String, azl1\textbar\textbar AZU)=0)
if Type == ``N'' then return (datatype (String) ==``NUM'') if Type ==
``S'' then return(symbol (String) ==`BAD') if Type == ``U'' then return
(verify (String, AZU)=0) if Type == ``W'' then do

/* It may not be a number. \emph{/ if DtypeOne(String) == `CHAR' then
return `0' /} It can be ``Whole'' even if originally in exponential
notation, provided it can be written as non-exponential. \emph{/ if
pos(`E',\#DatatypeResult)\textgreater0 then return `0' /} It won't be
``Whole'' if there is a non-zero after the decimal point. */
InFraction=`0'

do j = 1 to length (String) ce = substr(String,j,1) if
pos(c,`Ee')\textgreater0 then leave j if InFraction \&
pos(c,`+-')\textgreater0 then leave j if c == `.' then InFraction=`1'
else if InFraction \& a==`0' then return 0 end j /* All tests for Whole
passed. \emph{/ \#DatatypeResult = \#DatatypeResult \% 1 return 1 end /}
W \emph{/ /} Type will be ``x'' \emph{/ if pos (left
(String,1),\#Al1Blanks)\textgreater0 then return 0 if pos (right
(String,1),\#Al1Blanks)\textgreater0 then return 0 HexDigits=0 do
j=length(String) by -1 to 1 c=substr (String,j,1) if
pos(c,\#Al11Blanks)\textgreater0 then do /} Blanks need a pair of
HexDigits to the right of them. \emph{/ if HexDigits//2 = 0 then return
0 end else do if verify(c,``abcdefABCDEF''D09) = 0 then return 0
HexDigits=HexDigits+1 end end return 1 /} end X */

DtypeOne: /* See section nnn for the syntax of a number. \emph{/
\#DatatypeResult = `S' /} If not syntactically a number \emph{/ Residue
= strip(String) /} Blanks are allowed at both ends. */ if Residue ==
'\,' then return ``CHAR'' Sign = '\,' if left(Residue,1) == `+'
\textbar{} left(Residue,1) == `-' then do

Sign = left(Residue, 1) Residue = strip(substr(Residue,2),`L') /* Blanks
after sign */

end if Residue == '\,' then return ``CHAR'' /* Now testing Number,
section nnn */ if left(Residue,1) == `.' then do Residue = substr
(Residue, 2) Before = '\,' After = DigitRun() if After == '\,' then
return ``CHAR'' end else do Before = DigitRun() if Before == '\,' then
return ``CHAR'' if left(Residue,1) == `.' then do Residue = substr
(Residue, 2) After = DigitRun() end end Exponent = 0 if Residue == '\,'
then do if left(Residue, 1) == `e' \& left(Residue, 1) == `E' then

return ``CHAR'' Residue = substr (Residue, 2)

if Residue == '\,' then return ``CHAR''

Esign = '\,'

if left(Residue, 1) == `+' \textbar{} left(Residue, 1) == `-' then do
Esign = left(Residue, 1) Residue = substr (Residue, 2) if Residue ==
'\,' then return ``CHAR'' end

Exponent = DigitRun()

if Exponent == '\,' then return ``CHAR''

Exponent = Esign \textbar\textbar{} Exponent

end

if Residue == '\,' then return ``CHAR''

102 /\emph{DATATYPE tests for exponent out of range. }/

\#DatatypeResult = `E' /* If exponent out of range */ Before =
strip(Before,`L',`0') if Before == '\,' then Before = `0'

Exponent = Exponent + length(Before) -1 /* For SCIENTIFIC */

/* ``Engineering notation causes powers of ten to expressed as a
multiple of 3 - the integer part may therefore range from 1 through
9910.'' \emph{/ g=1 if \#Form.\#Level == `E' then do /} Adjustment to
make exponent a multiple of 3 */ g = Exponent//3 if g \textless{} 0 then
g = g + Exponent = Exponent - end

/* Check on the exponent. */ if Exponent \textgreater{} \#Limit
ExponentDigits then return ``CHAR'' if -\#Limit ExponentDigits
\textgreater{} Exponent then return ``CHAR''

/* Format to the numeric setting of the caller of DATATYPE */ numeric
digits \#Digits.\#Level

numeric form value \#Form.\#Level

\#DatatypeResult = 0 + \#Bif\_Arg.1

return ``NUM''

DigitRun: Outcome = '\,' do while Residue == '\,' if pos(left (Residue,
1), `0123456789') = 0 then leave

Outcome = Outcome \textbar\textbar{} left(Residue, 1) Residue =
substr(Residue, 2) end

return Outcome

@@

call CheckArgs `rANY rWHOLE\textgreater0 oWHOLE\textgreater=0' String
\#Bif Arg.1

Num \#Bif Arg.2 if \#Bif\_ArgExists.3 then Len = \#Bif\_Arg.3

if Num \textgreater{} length(String) then return String

Output = substr(String, 1, Num - 1) if \#Bif\_ArgExists.3 then if Num +
Len \textless= length(String) then Output = Output \textbar\textbar{}
substr(String, Num + Len) return Output

@@

call CheckArgs `rANY rWHOLE\textgreater0 oWHOLE\textgreater=0' String
\#Bif Arg.1

Num \#Bif Arg.2 if \#Bif\_ArgExists.3 then Len = \#Bif\_Arg.3

if Num \textgreater{} words(String) then return String

EndLeft = wordindex(String, Num) - 1 Output = left(String, EndLeft) if
\#Bif\_ArgExists.3 then do BeginRight = wordindex(String, Num + Len)

if BeginRight\textgreater0 then Output =

end return Output

@@

call CheckArgs

New \#Bif\_ Arg.1 Target \#Bif Arg.2 if \#Bif\_ArgExists.3

then else then else then else

Num Num Length = \#Bif Arg.4 Length = length (New) Pad \#Bif Arg.5 Pad
ro

\#Bif Arg.3 0

if \#Bif\_ArgExists.4

if \#Bif\_ArgExists.5

return left(Target, Num, Pad) left (New, Length, Pad), substr(Target,
Num + 1)

``

@@

`rANY rANY oWHOLE\textgreater=0 oWHOLE\textgreater=0 oPAD'

\emph{/ }/

LASTPOS returns the position of the last occurrence of the first
argument within the second. The third

argument is a starting position for the search.

call CheckArgs `rANY rANY oWHOLE\textgreater0'

Needle = \#Bif\_Arg.1 Haystack = \#Bif Arg.2 if \#Bif\_ArgExists.3 then
Start = \#Bif\_Arg.3 else Start = length(Haystack)

NeedleLength = length (Needle) if NeedleLength = 0 then return 0 Start =
Start - NeedleLength + 1 do i= Start by -1 while i \textgreater{} 0 if
substr(Haystack, i, NeedleLength) end i return 0

@@

length of the result and the third is the padding character. call
CheckArgs `rANY rWHOLE\textgreater=0 oPAD'

if \#Bif\_ArgExists.3 then Pad else Pad

\#Bif Arg.3

return substr(\#Bif Arg.1, 1, \#Bif\_ Arg.2, Pad)

10.1.14 LENGTH

Needle then return i

The second argument specifies the

Length returns a count of the number of characters in the argument.

call CheckArgs `rANY'

String = \#Bif Arg.1

\#Response = Config Length(String)

Length = \#Outcome

call Config Substr \#Response, 1

if \#Outcome == `E' then return Length

/* Here if argument was not a character string. call Config C2B String
call \#Raise `SYNTAX',

23.1, b2x(\#Outcome)

\emph{/ /} No return to here */

@@

OVERLAY overlays the first argument onto the second. The third argument
is the starting position of the

overlay. The fourth argument is the length of the overlay and the fifth
is the padding character.

call CheckArgs `rANY rANY oWHOLE\textgreater0 oOWHOLE\textgreater=0
oPAD'

New = \#Bif\_Arg.1

Target = \#Bif Arg.2

if \#Bif\_ArgExists.3 then Num = \#Bif\_Arg.3 else Num = 1

if \#Bif\_ArgExists.4 then Length = \#Bif\_ Arg.4 else Length = length
(New)

if \#Bif\_ArgExists.5 then Pad = \#Bif\_ Arg.5 else Pad = ' ' return
left(Target, Num - 1, Pad), /* To left of overlay \emph{/
\textbar\textbar{} left (New, Length, Pad), /} New string overlaid
\emph{/ \textbar\textbar{} substr (Target, Num + Length) /} To right of
overlay */

10.1.16 POS POS returns the position of the first argument within the
second.

call CheckArgs `rANY rANY oWHOLE\textgreater0'

Needle \#Bif Arg.1

Haystack = \#Bif Arg.2

if \#Bif\_ ArgExists.3 then Start else Start

\#Bif Arg.3 1

if length(Needle) = 0 then return 0

do i = Start to length (Haystack) +1-length (Needle) if substr(Haystack,
i, length(Needle)) == Needle then return i end i

return 0

@@

call CheckArgs `rANY'

String

\#Bif Arg.1 Output = '\,' do i= 1 to length (String) Output =
substr(String,i,1) \textbar\textbar{} Output end return Output

@@

length of the result and the third is the padding character. call
CheckArgs `rANY rWHOLE\textgreater=0 oPAD'

String = \#Bif Arg.1 Length = \#Bif Arg.2 if \#Bif\_ArgExists.3 then Pad

\#Bif Arg.3 else Pad '

Trim = length(String) - Length if Trim \textgreater= 0 then return
substr(String,Trim + 1) return copies(Pad, -Trim) \textbar\textbar{}
String /* Pad string on the left */

@@

call CheckArgs `rANY oOWHOLE\textgreater=0 oPAD'

String = \#Bif Arg.1

if \#Bif\_ArgExists.2 then Num = \#Bif\_ Arg.2 else Num = 1 if \#Bif
ArgExists.3 then Pad = \#Bif Arg.3 else Pad = ' '! Padding = copies(Pad,
Num) Output = subword(String, 1, 1) do i = 2 to words (String) Output =
Output \textbar\textbar{} Padding \textbar\textbar{} subword(String, i,
1)

end return Output

@@

call CheckArgs `rANY oLTB oPAD'

String = \#Bif Arg.1 if \#Bif\_ArgExists.2 then Option = \#Bif Arg.2
else Option = `B' if \#Bif\_ ArgExists.3 then Unwanted = \#Bif\_ Arg.3
else Unwanted = \#A11Blanks\textless Index ``\#Al11Blanks'' \# ``\,''
\textgreater{}

if Option == `L' \textbar{} Option == `B' then do /* Strip leading
characters */ do while String == '\,' \& pos(left(String, 1), Unwanted)
\textgreater{} 0 String = substr(String, 2) end end

if Option == `T' \textbar{} Option == `B' then do /* Strip trailing
characters \emph{/ do while String == '\,' \& pos(right(String, 1),
Unwanted) \textgreater{} 0 String = left(String, length(String) -1) end
/} of while */ end return String

@@

SUBSTR returns a sub-string of the first argument. The second argument
specifies the position of the first character and the third specifies
the length of the sub-string. The fourth argument is the padding

character.

call CheckArgs `rANY rWHOLE\textgreater0 oOWHOLE\textgreater=0 oPAD'

String = \#Bif Arg.1 Num = \#Bif\_ Arg.2 if \#Bif\_ArgExists.3 then
Length = \#Bif Arg.3 else Length = max(length (String) +1-Num, 0) if
\#Bif\_ArgExists.4 then Pad = \#Bif\_ Arg.4 else Pad = ' `! Output =''
do Length \#Response Config Substr(String,Num) /* Attempt to fetch
character.*/ Character \#Outcome

Num = Num + 1 call Config Substr \#Response,1 /* Was there such a
character? */ if \#Outcome == `E' then do

/* Here if argument was not a character string. */

call Config C2B String call \#Raise `SYNTAX', 23.1, b2x(\#Outcome) /* No
return to here */

end if \#Outcome == `M' then Character = Pad Output=Output \textbar{}
\textbar{} Character end

return Output

@@

call CheckArgs `rANY rWHOLE\textgreater0 oWHOLE\textgreater=0'

String \#Bif Arg.1

Num \#Bif Arg.2

if \#Bif\_ ArgExists.3 then Length else Length

\#Bif\_Arg.3 length(String) /* Avoids call \emph{/ /} to WORDS() */

if Length = 0 then return '\,'

/* Find position of first included word */

Start = wordindex (String, Num)

if Start = 0 then return '\,' /* Start is beyond end */

/* Find position of first excluded word */ End = wordindex (String,
Num+Length) if End = 0 then End = length(String)+1

Output=substr (String, Start, End-Start)

/* Drop trailing blanks */

do while Output == '\,' if pos (right (Output,1),\#A11Blanks) = 0 then
leave Output = left (Output, length (Output) -1) end

return Output

@@

call CheckArgs `rANY oANY oANY oPAD' String = \#Bif Arg.1 /* If neither
input nor output tables, uppercase. */ if \#Bif ArgExists.2 \&
\#Bif\_ArgExists.3 then do Output = '\,' do j=1 to length (String)
\#Response = Config Upper (substr (String,j,1))

Output = Output \textbar\textbar{} \#Outcome end j

return Output

end

/* The input table defaults to all characters. \emph{/ if \#Bif
ArgExists.3 then do \#Response = Config Xrange() Tablei = \#Outcome end
else Tablei = \#Bif\_Arg.3 /} The output table defaults to null \emph{/
if \#Bif\_ArgExists.2 then Tableo = \#Bif Arg.2 else Tableo = '\,' /}
The tables are made the same length */ if \#Bif\_ArgExists.4 then Pad =
\#Bif\_ Arg.4 else Pad = ' ' Tableo=left (Tableo, length (Tablei) , Pad)

107 Output='\,' do j=1 to length (String) c=substr (String,j,1)
k=pos(c,Tablei) if k=0 then Output=Output \textbar\textbar c else
Output=Output \textbar{} \textbar{} substr (Tableo,k,1) end j return
Output

@@

call CheckArgs `rANY rANY oMN oWHOLE\textgreater0'

String = \#Bif Arg.1

Reference = \#Bif\_Arg.2

if \#Bif\_ ArgExists.3 then Option \#Bif Arg.3 else Option 'N!

if \#Bif\_ ArgExists.4 then Start = \#Bif\_Arg.4 else Start = 1

Last = length(String) if Start \textgreater{} Last then return 0 if
Reference == '\,' then if Option == `N' then return Start else return 0

do i = Start to Last t = pos(substr(String, i, 1), Reference) if Option
== `N' then do if t = 0 then return i /* Return position of NoMatch
character. \emph{/ end else if t \textgreater{} 0 then return i /}
Return position of Matched character. */ end i return 0

@@

call CheckArgs `rANY rwWHOLE\textgreater0'

return subword(\#Bif Arg.1, \#Bif\_Arg.2, 1)

@@

call CheckArgs `rANY rwWHOLE\textgreater0'

String Num

\#Bif Arg.1 \#Bif Arg.2

/* Find starting position */

Start = 1 Count = 0 do forever Start = verify(String,
\#Al1Blanks\textless Index ``\#A11Blanks'' \# ``\,'' \textgreater, `N',
Start) /\emph{ Find non-blank }/ if Start = 0 then return 0 /* Start is
beyond end \emph{/ Count = Count + 1 /} Words found \emph{/ if Count =
Num then leave Start = verify(String, \#Al1Blanks\textless Index
``\#Al11Blanks'' \# ``\,'' \textgreater, `M', Start + 1) /}

Find blank */

if Start = 0 then return 0 /* Start is beyond end */ end return Start

@@

call CheckArgs `rANY rwWHOLE\textgreater0'

return length(subword(\#Bif Arg.1, \#Bif\_Arg.2, 1))

@@

call CheckArgs `rANY rANY oWHOLE\textgreater0'

Phrase = \#Bif\_Arg.1

String = \#Bif Arg.2

>>>>

\lstinputlisting[language=rexx,label=argumentchecking.rexx,caption=argumentchecking.rexx]{argumentchecking.rexx}

10.2.2 Date calculations

\lstinputlisting[language=rexx,label=datecalculations.rexx,caption=datecalculations.rexx]{datecalculations.rexx}

@@

\lstinputlisting[language=rexx,label=reradix.rexx,caption=reradix.rexx]{reradix.rexx}

@@

\lstinputlisting[language=rexx,label=raise.rexx,caption=raise.rexx]{raise.rexx}

@@

\lstinputlisting[language=rexx,label=abbrev.rexx,caption=abbrev.rexx]{abbrev.rexx}

@@

\lstinputlisting[language=rexx,label=center.rexx,caption=center.rexx]{center.rexx}

@@

\lstinputlisting[language=rexx,label=changestr.rexx,caption=changestr.rexx]{changestr.rexx}

@@

\lstinputlisting[language=rexx,label=compare.rexx,caption=compare.rexx]{compare.rexx}

@@

\lstinputlisting[language=rexx,label=copies.rexx,caption=copies.rexx]{copies.rexx}

@@

\lstinputlisting[language=rexx,label=counstr.rexx,caption=counstr.rexx]{counstr.rexx}

@@

\lstinputlisting[language=rexx,label=datatype.rexx,caption=datatype.rexx]{datatype.rexx}

@@

\lstinputlisting[language=rexx,label=delstr.rexx,caption=delstr.rexx]{delstr.rexx}

@@

\lstinputlisting[language=rexx,label=delword.rexx,caption=delword.rexx]{delword.rexx}

@@

\lstinputlisting[language=rexx,label=insert.rexx,caption=insert.rexx]{insert.rexx}

@@

\lstinputlisting[language=rexx,label=lastpos.rexx,caption=lastpos.rexx]{lastpos.rexx}

@@

length of the result and the third is the padding character.

\lstinputlisting[language=rexx,label=left.rexx,caption=left.rexx]{left.rexx}

@@

OVERLAY overlays the first argument onto the second. The third argument
is the starting position of the overlay. The fourth argument is the
length of the overlay and the fifth is the padding character.

\lstinputlisting[language=rexx,label=overlay.rexx,caption=overlay.rexx]{overlay.rexx}

@@

\lstinputlisting[language=rexx,label=reverse,caption=reverse]{reverse}

@@

length of the result and the third is the padding character.

\lstinputlisting[language=rexx,label=right.rexx,caption=right.rexx]{right.rexx}

@@

\lstinputlisting[language=rexx,label=space.rexx,caption=space.rexx]{space.rexx}

@@

\lstinputlisting[language=rexx,label=strip.rexx,caption=strip.rexx]{strip.rexx}

@@

SUBSTR returns a sub-string of the first argument. The second argument
specifies the position of the first character and the third specifies
the length of the sub-string. The fourth argument is the padding
character.

\lstinputlisting[language=rexx,label=substr.rexx,caption=substr.rexx]{substr.rexx}

@@

\lstinputlisting[language=rexx,label=subword.rexx,caption=subword.rexx]{subword.rexx}

@@

\lstinputlisting[language=rexx,label=translate,caption=translate]{translate}

@@

\lstinputlisting[language=rexx,label=verify.rexx,caption=verify.rexx]{verify.rexx}

@@

\lstinputlisting[language=rexx,label=word.rexx,caption=word.rexx]{word.rexx}

@@

\lstinputlisting[language=rexx,label=wordindex.rexx,caption=wordindex.rexx]{wordindex.rexx}

@@

\lstinputlisting[language=rexx,label=wordlength.rexx,caption=wordlength.rexx]{wordlength.rexx}

@@

\lstinputlisting[language=rexx,label=wordpos.rexx,caption=wordpos.rexx]{wordpos.rexx}
